### 2024-04-22 docs/SOSP 2024/exploring.tex
Final submission

% \begin{itemize}
%   \item The scheduling language described in Section \ref{sec:schedule} allows
%   for numerous possible schedules for a given model. 
%   Finding a schedule with good performance is a non-trivial task.
%   \item We identify a reasonable template schedule for GPUs which encompasses
%   all strategies published in prior work.
%   \item Describe the template schedule.
%   \item Even within the variants of this template schedule, there is a significant 
%   amount of variation in performance. Show the histogram. 
%   \item Extensively exploring all possible parameter values for the template schedule 
%   is infeasible due to the large number of parameters.
%   Cite some numbers on how long it takes to explore the schedule space.
%   \item List the observations we make and the heuristic we design based on these.
% \end{itemize}
The set of schedules that can be constructed using the scheduling language described in 
Section \ref{sec:schedule} is unbounded.
%\TODO{kr: next few lines need more work}
% Even if one were to search for a good schedule offline, before deploying the model 
% on a specific hardware, additional tools are needed to help search for a 
% high-performance schedule. 
Additional tools are required to help search for a high-performance schedule.
We anticipate that even for the most complex models, 
% one would like to spend at most a few minutes doing this search.
spending more than a few minutes on this search is not acceptable.
We propose a set of heuristics to guide the search for a good schedule 
and meet this goal. These heuristics work by first defining a 
bounded search space and then pruning this search space further
based on the batch size and model characteristics.

\subsection{Bounding the search space}
We bound the space using a few meta-parameters that together define 
the primitives used to construct a schedule. We do so while making sure 
that the space (at-least) covers the known strategies published in prior work.
Specifically, we use 4 numeric parameters, 4 boolean parameters and one ternary parameter as listed in Table \ref{tab:schedparams}.
We picked the specific values listed through experimentation.
%Note that even this template has a large number ($1728$) of schedules, and requires further pruning as we discuss later.
The first three numeric parameters assign a configurable number of rows to each 
thread block, to each thread, and determine how trees are distributed across a
specified number of threads. These parameters together define the loop schedule 
primitives to use (including the arguments to pass) from Table~\ref{Tab:LoopModifiers} 
and the parallelization strategy. The next four parameters determine the caching strategy, 
unroll factor{\footnote{We choose to always unroll trees completely, while its possible 
to partially unroll loops we did not observe any performance benefits from doing so.}} 
and how many trees to traverse simultaneously within
a single thread (the remaining primitives from Table~\ref{Tab:Optimizations}).  
The last two parameters determine reduction optimizations (Table~\ref{Tab:ReductionOpts}) 
and the layout{\footnote{This does not have a schedule primitive. We always explore all three layout options.}} to use.
%Table \ref{tab:schedparams} lists the parameter values we tried. 
% \begin{itemize}
%   \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
%   \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
%   \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
%   \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
%   \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
%   \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
%   \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
%   \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
% \end{itemize}

>>>>

% \begin{itemize}
%   \item The scheduling language described in Section \ref{sec:schedule} allows
%   for numerous possible schedules for a given model. 
%   Finding a schedule with good performance is a non-trivial task.
%   \item We identify a reasonable template schedule for GPUs which encompasses
%   all strategies published in prior work.
%   \item Describe the template schedule.
%   \item Even within the variants of this template schedule, there is a significant 
%   amount of variation in performance. Show the histogram. 
%   \item Extensively exploring all possible parameter values for the template schedule 
%   is infeasible due to the large number of parameters.
%   Cite some numbers on how long it takes to explore the schedule space.
%   \item List the observations we make and the heuristic we design based on these.
% \end{itemize}
The set of schedules that can be constructed using the scheduling language described in 
Section \ref{sec:schedule} is unbounded.
%\TODO{kr: next few lines need more work}
% Even if one were to search for a good schedule offline, before deploying the model 
% on a specific hardware, additional tools are needed to help search for a 
% high-performance schedule. 
Additional tools are required to help search for a high-performance schedule.
Like with other schedule-guided ML compilers~\cite{TVM},
we expect schedule exploration will be performed offline (before a model deployment),
% But still, even for the most complex models, 
% one would like to spend at most a few minutes doing this search.
but spending more than a few minutes on this search is likely not acceptable.
We propose a set of heuristics to guide the search for a good schedule 
and meet this goal. These heuristics work by first defining a 
bounded search space and then pruning this search space further
based on the batch size and model characteristics.

\subsection{Bounding the search space}
We bound the space using a few meta-parameters that together define 
the primitives used to construct a schedule. We do so while making sure 
that the space (at-least) covers the known strategies published in prior work.
Specifically, we use the 9 parameters listed in Table \ref{tab:schedparams}.
We picked the specific values listed through experimentation.
%Note that even this template has a large number ($1728$) of schedules, and requires further pruning as we discuss later.
The first three numeric parameters assign a configurable number of rows to each 
thread block, to each thread, and determine how trees are distributed across a
specified number of threads. These parameters together define the loop schedule 
primitives to use (including the arguments to pass) from Table~\ref{Tab:LoopModifiers} 
and the parallelization strategy. The next four parameters determine the caching strategy, 
unroll factor{\footnote{We choose to always unroll trees completely, while its possible 
to partially unroll loops we did not observe any performance benefits from doing so.}} 
and how many trees to traverse simultaneously within
a single thread (the remaining primitives from Table~\ref{Tab:Optimizations}).  
The last two parameters determine reduction optimizations (Table~\ref{Tab:ReductionOpts}) 
and the layout{\footnote{This does not have a schedule primitive. We always explore all three layout options.}} to use.
%Table \ref{tab:schedparams} lists the parameter values we tried. 
% \begin{itemize}
%   \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
%   \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
%   \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
%   \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
%   \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
%   \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
%   \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
%   \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
% \end{itemize}

### 2024-04-22 docs/SOSP 2024/related.tex
Minor edits
Final submission

On CPUs, XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and
scikit-learn\cite{Sklearn} are extremely popular.
A recent paper~\cite{PACTVanLunteren} describes an adaptive mechanism 
to pick one of a few predefined parallelization and vectorization strategies.
Other systems that hide dependency stalls by interleaving tree walks\cite{VPred},
implement optimized algorithms for tree inference\cite{QuickScorer, QuickScorer1}
and improve cache performance of decision tree ensembles on CPUs\cite{CacheConscious1, CacheConscious2}
have been proposed in prior work.
Some systems have been proposed to parallelize decision tree training 
on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.
However, none of these systems provide portable performance across different target machines.
While the specifics may vary, \Treebeard{} supports compiler optimizations that 
achieve similar results as techniques in these systems.

@@

\emph{Reductions:} CUB\cite{CUB} and Thrust\cite{Thrust} are libraries 
that implement high-performance parallel reductions on GPUs. 
However, it is not possible to fuse these functions with other computations  
as required in \Treebeard{}. Reddy et. al.~\cite{ChandanReduction}
describe language constructs in {\sc {Pencil}}~\cite{Pencil}
to express reductions and to represent and optimize them using the polyhedral 
framework. 
% It is not clear how these techniques can be fused with 
% other computations in arbitrary loop nests as required in \Treebeard{}.
%Additionally, 
Their system does not express the hierarchical nature of 
reductions and also only targets GPUs. Suriana et. al.~\cite{HalideReductions}
extend Halide to add support for factoring reductions in the Halide 
scheduling language and to synthesize reduction operators. De Gonzalo 
et. al.~\cite{TangramReduction} describe a system based on Tangram 
that composes several partial reduction implementations into different 
reduction implementations for GPUs and then searches through these
alternate implementations to find the best ones. In summary, none of 
these systems provide abstractions and a general framework to generate 
and optimize reductions across different target processors as \Treebeard{} does.

>>>>

On CPUs, XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and
scikit-learn\cite{Sklearn} are extremely popular.
A recent paper~\cite{PACTVanLunteren} describes an adaptive mechanism 
to pick one of a few predefined parallelization and vectorization strategies.
Other systems that hide dependency stalls by interleaving tree walks\cite{VPred},
implement optimized algorithms for tree inference\cite{QuickScorer, QuickScorer1}
and improve cache performance of decision tree ensembles on CPUs\cite{CacheConscious1, CacheConscious2}
have been proposed.
Some systems have been proposed to parallelize decision tree training 
on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.
However, none of these systems provide portable performance across different target machines.
While the specifics may vary, \Treebeard{} supports retargetable compiler optimizations that 
achieve similar results. % as techniques in these systems.

@@

\emph{Reductions:} CUB\cite{CUB} and Thrust\cite{Thrust} are libraries 
that implement reductions on GPUs. 
However, it is not possible to fuse these functions with other computations  
as required in \Treebeard{}. Reddy et. al.~\cite{ChandanReduction}
describe language constructs in {\sc {Pencil}}~\cite{Pencil}
to express reductions and optimize them using the polyhedral 
framework. 
% It is not clear how these techniques can be fused with 
% other computations in arbitrary loop nests as required in \Treebeard{}.
%Additionally, 
Their system does not express the hierarchical nature of 
reductions and also only targets GPUs. Suriana et. al.~\cite{HalideReductions}
extend Halide to add support for reductions in the Halide 
scheduling language and to synthesize reduction operators. De Gonzalo 
et. al.~\cite{TangramReduction} describe a system based on Tangram 
that composes several partial reduction implementations into different 
reduction implementations for GPUs and then searches through these
alternate implementations to find the best ones. In summary, none of 
these systems provide abstractions and a general framework to generate 
and optimize reductions across different target processors as \Treebeard{} does.
